---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GAclustenhancer

<!-- badges: start -->
<!-- badges: end -->

The goal of GAclustenhancer is to use genetic algorithms to improve clustering in situations where clusters are used to calculate correlation. A GA with a correlation-based fitness function optimizes clustering results to maximize correlation values. A common use for this package is with gene expression data analysis. An example using sample gene expression data is shown below.

This package requires a few imports: GA, methods, Rfast, stats, utils

While not necessary, we suggest you also import DESeq2, SummarizedExperiment for working with gene datasets.

## Installation

You can install the development version of GAclustenhancer from [GitHub](https://github.com/frankcsquared/GAclustenhancer) with:

``` r
# install.packages("devtools")
devtools::install_github("frankcsquared/GAclustenhancer")
```

## Example

This is a basic example using sample data which shows you how to solve a common problem:

```{r example}
## load in packages
library(GAclustenhancer)

## set paths and read in data (change paths before using)
load("./data/count_data.rda") #gene expression count data of dim 882 genes by 348 cell line samples, preprocessed in data-raw/preprocess.R
load("./data/t_count_data.rda") #transposed and preprocessed gene expression count data
load("./data/input_lfc.rda") #L2fc results calculated from DESeq2

## run example on subset of dataset
obj_gene <- ga.clust(dataset = t_count_data[1:100, 1:50], k = 2, plot.internals = TRUE, seed.p = 42, pop.size = ncol(count_data[1:50, 1:100]), known_lfc = input_lfc[1:50])

## run example on entire dataset (runtime > 1 hour)
#obj_gene <- ga.clust(dataset = t_data, k = 2, plot.internals = TRUE, seed.p = 42, pop.size = ncol(data), known_lfc = input_lfc)
```

## Comparison with Non-GA Enhanced Clustering

Our GA-based clustering algorithm acheives a correlation of 0.826 by iteratively optimizing the cluster partition until high correlation is achieved. Non-GA enhanced clustering (hclust function in R) acheives a correlation of 0.693 on the same dataset, as shown below:

```{r}
library(GAclustenhancer)

## set paths and read in data (change paths before using)
load("./data/count_data.rda") #gene expression count data of dim 882 genes by 348 cell line samples, preprocessed in data-raw/preprocess.R
load("./data/t_count_data.rda") #transposed and preprocessed gene expression count data
load("./data/input_lfc.rda") #L2fc results calculated from DESeq2

## cluster data
cluster_results <- cluster.l2fc(count_data, input_lfc)

## correlate with existing l2fc
abs(cor(cluster_results$lfc, input_lfc))
```
